Using Metric Space Indexing for Complete and Efficient Record Linkage

### 2017-11-21 paper.tex
Update on Overleaf.

% ====================================================================

@@

\fi
% ====================================================================

@@

@@

% --------------------------------------------------------------------

@@

We now describe the datasets we use in our evaluation, the set-up
employed to evaluate our proposed metric indexing approach and compare
it with traditional blocking techniques, and we then present and discuss
our results~\footnote{The raw data, additional figures and the source
code required to run these experiments can be downloaded from here.
(Link removed for double blind review)}.

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

@@

The other datasets are historical Scottish records of vital events
(birth, marriages and deaths) one registered on the \emph{Isle of Skye},
a rural district, and the other records from \emph{Kilmarnock}, an
industrial town. These datasets were create by historical demographers
who extensively curated and linked both
datasets~\cite{reid2002,reid2006}. Both data sets include the names,
gender, addresses of individuals and their parents. Ground truth was
generated by the demographers based on their extensive domain knowledge.

@@

\subsection{Cora}

In this section, we perform linkage on the Cora dataset using all four
approaches presented in this paper: brute force, traditional blocking,
LSH and M-tree. The brute-force approach is used to establish the
threshold value that yields the best possible linkage for the Cora
dataset. The threshold value is varied between 0 and 250,
\Cref{cora-quality} presents the precision, recall, and F-measure
values. The linkage results have high precision and low recall for low
threshold values, and high recall and low precision for high threshold
values. For this dataset, the maximum F-measure value is achieved using
a threshold value of 70. This value is data dependent; for different
datasets the maximum F-measure will correspond to different thresholds.
In the rest of this section we fix the threshold value to be 70 which
maximises the F-measure for this dataset.

@@

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-authors}
\caption{Cora - Traditional Blocking on the `authors' field}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-title}
\caption{Cora - Traditional Blocking on the `title' field}
\end{subfigure}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-year}
\caption{Cora - Traditional Blocking on the `publication year' field}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-combined}
\caption{Cora - Traditional Blocking using the union of blocking on all fields}
\end{subfigure}

@@

\caption{Linkage results on the Cora data set}
\label{cora-quality}
\end{figure}

% \begin{figure}
% \includegraphics[width=\textwidth]{figures/plotLQ-cora-brute}
% \caption{Linkage quality using brute-force linking on the Cora dataset\label{fig-cora-brute}}
% \end{figure}

In the traditional blocking experiment the following fields from the
Cora dataset are used individually as blocking keys: \textit{author},
\textit{title}, \textit{venue}, \textit{location}, \textit{publisher}
and \textit{year}. We also use a combined blocking key comprising all
fields.

\Cref{comparison-of-results-cora} shows greater detail for a subset of
the Cora experiments, with the distance threshold fixed at 70. The table
shows the parameters for the experiment, the number of distance
comparisons made, the precision, recall and the F1 measure achieved by
each algorithm.  In the \emph{Linker} column the algorithm name is
followed by its parameters: for LSH the number of the bands followed by
the band size, and for traditional blocking the fields used for
blocking. The number of distance comparisons is reported as a
machine-independent proxy for execution cost, since code profiling shows
that distance calculations are dominant.

M-tree yields the same linkage quality as brute-force, although using a significantly lower
number of comparisons. This is as expected, since both techniques are complete.

Three of the LSH configurations give similar quality

\begin{table}[ht]
\centering
\begin{tabular}{l|r|r|r|r}
\parbox{0.2\textwidth}{\centering Linker} &
\parbox{0.2\textwidth}{\centering Comparisons} &
\parbox{0.17\textwidth}{\centering Precision} &
\parbox{0.17\textwidth}{\centering Recall} &
\parbox{0.17\textwidth}{\centering F1 Measure} \\ \hline \hline
Brute Force        & 1,677,025                 & 0.84      & 0.57   & 0.68       \\ \hline
MTree              &  902,693                  & 0.84      & 0.57   & 0.68       \\ \hline
LSH-2-2            &  192,199	& 0.95	& 0.47	& 0.63       \\
LSH-5-2            &  342,849	& 0.91	& 0.55	& 0.69       \\
LSH-10-2           &  513,947	& 0.88	& 0.57	& 0.69       \\
LSH-2-5            &  14,329	& 0.99	& 0.28	& 0.43      \\
LSH-5-5            &  22,057	& 0.99	& 0.36	& 0.53       \\
LSH-10-5           &  26,167	& 0.98	& 0.4	& 0.57      \\
LSH-2-10           &  4,711	& 1	& 0.15	& 0.27    \\
LSH-5-10           &  6,501	& 1	& 0.19	& 0.32      \\
LSH-10-10          & 10,627  	& 0.99	& 0.27& 	0.43       \\  \hline
Block-year         &  115,893                  & 0.99      & 0.35   & 0.51       \\
Block-authors      &   11,039                   & 0.94      & 0.16   & 0.28       \\
Block-title        &   27,407                   & 0.95      & 0.42   & 0.58       \\
Block-venue        &   36,647                   & 0.85      & 0.29   & 0.44       \\
Block-location     & 1,009,957                 & 0.83      & 0.43   & 0.57       \\
Block-publisher    &  833,079                  & 0.85      & 0.44   & 0.58       \\
Block-combined     & 1,214,269                 & 0.84      & 0.56   & 0.67       \\ \hline
\end{tabular}
\caption{Linkage quality on Cora dataset with distance threshold = 70
\label{comparison-of-results-cora}
}
\end{table}

\subsection{Historical Demography results}

@@

\begin{table}[ht]
\centering
\begin{tabular}{l|r|r|r|r}
\parbox{0.2\textwidth}{\centering Linker} &
\parbox{0.2\textwidth}{\centering Comparisons} &
\parbox{0.17\textwidth}{\centering Precision} &
\parbox{0.17\textwidth}{\centering Recall} &
\parbox{0.17\textwidth}{\centering F1 Measure} \\ \hline \hline
MTree    & 102,318,525                 & 0.65      & 0.46   & 0.54      \\ \hline

@@

%-------------------------------------------------------------------------------------

\begin{table}[ht]
\centering
\begin{tabular}{l|r|r|r|r|r|r}
\parbox{0.2\textwidth}{\centering Linker} &
\parbox{0.2\textwidth}{\centering Comparisons} &
\parbox{0.17\textwidth}{\centering Precision} &
\parbox{0.17\textwidth}{\centering Recall} &
\parbox{0.17\textwidth}{\centering F1 Measure} \\ \hline \hline
MTree     & 514,871,153                   & 0.76      & 0.45   & 0.57      \\ \hline
LSH-2-2   & 99,145,887                    & 0.81      & 0.16   & 0.27       \\
LSH-5-2   & 130,721,338                   & 0.79      & 0.23   & 0.36       \\
LSH-10-2  & 177,168,848                   & 0.79      & 0.36   & 0.49       \\
LSH-5-5   & 239,368                      & 0.84      & 0.01   & 0.02       \\
LSH-10-5  & 855,431                      & 0.87      & 0.02   & 0.03 \\ \hline
\end{tabular}
\caption{Linkage quality on Kilmarnock dataset with distance threshold = 2
\label{comparison-of-results-demography-kili}
}
\end{table}

%-------------------------------------------------------------------------------------

@@

However, this paper also highlights some unexpected results. Firstly, the good quality results yielded by both traditional blocking and LSH are partly due to the fact that (in the limit) they tend towards brute-force as the number of records in the block increase. A second unexpected result is that shown in \Cref{quality-cora}, namely that LSH can yield higher precision than is achieved by a complete method such as M-tree, as was discussed in Sect.~\ref{sec-approach}.

>>>>

%----------------------------------------------------------------------

@@

\fi
%----------------------------------------------------------------------

@@

%----------------------------------------------------------------------

@@

%----------------------------------------------------------------------

@@

We now describe the datasets we use in our evaluation, the set-up
employed to evaluate our proposed metric indexing approach and compare
it with traditional blocking techniques, and we then present and discuss
our results~\footnote{The raw data, additional figures and the source
code required to run these experiments can be downloaded from:
(link removed for double blind review).}.

%----------------------------------------------------------------------

@@

The other datasets are historical Scottish records of vital events
(birth, marriages and deaths) one registered on the \emph{Isle of Skye},
a rural district, and the other records from \emph{Kilmarnock}, an
industrial town. These datasets were create by historical demographers
who extensively curated and linked both
datasets~\cite{reid2002,reid2006}. Both datasets include the names,
gender, addresses of individuals and their parents. Ground truth was
generated by the demographers based on their extensive domain knowledge.

@@

\subsection{Cora Results}

In this section, we perform linkage on the Cora dataset using all four
approaches presented in this paper: brute force, traditional blocking,
LSH and M-tree, using several selected configurations for the blocking
and LSH linkers. The distance threshold is varied between 0 and
250~\footnote{Relatively high Levenshtein distances are included since
Cora includes a number of low-similarity true matches.}. For traditional
blocking, the following Cora fields are used individually as blocking
keys: \textit{author}, \textit{title}, \textit{venue},
\textit{location}, \textit{publisher} and \textit{year}. We also use a
combined blocking key comprising all fields.

\Cref{cora-quality} shows the precision, recall, and F-measure
for various thresholds. As expected, low thresholds give high precision and low recall, and the reverse for high threshold
values. As expected, brute force and

For a more detailed investigation of selected linkers, the brute-force
approach is used to establish the optimum threshold value for the Cora
dataset. The maximum F-measure value is observed at a threshold value of
70. This value is dataset-dependent; for different datasets the maximum
F-measure will occur at different thresholds. In the rest of this
section we fix the threshold value at 70.

\Cref{comparison-of-results-cora} shows greater detail for selected
linkers, with the distance threshold fixed at 70. The table shows the
parameters for the experiment, the number of distance comparisons made,
the precision, recall and the F1 measure achieved by each algorithm.  In
the \emph{Linker} column the algorithm name is followed by its
parameters: for LSH the number of the bands followed by the band size,
and for traditional blocking the fields used for blocking. The number of
distance comparisons is reported as a machine-independent proxy for
execution cost, since code profiling shows that distance calculations
are dominant.

M-tree yields the same linkage quality as brute-force, although using a
significantly lower number of comparisons. This is as expected, since
both techniques are complete. Several of the incomplete linkers give
similar quality, for example \emph{LSH-2-2}, \emph{LSH-5-2}, \emph{LSH-10-2} and
\emph{Block-combined}. These, and a number of other incomplete linkers give
better precision than the complete techniques. This is due to
high-similarity non-matches, as discussed in Sect.~\ref{sec-approach}.
Although several of the incomplete linkers give as good quality as
M-tree, and in some cases at lower cost, this is offset by the need to
select appropriate configuration parameters. Some other linkers give
very poor results.

%----------------------------------------------------------------------

@@

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-authors}
\caption{Cora - Traditional Blocking on `authors' field}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-title}
\caption{Cora - Traditional Blocking on `title' field}
\end{subfigure}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-year}
\caption{Cora - Traditional Blocking on `publication year' field}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-combined}
\caption{Cora - Traditional Blocking using the union of blocking on all fields}
\end{subfigure}

@@

\caption{Linkage results on Cora dataset}
\label{cora-quality}
\end{figure}

%----------------------------------------------------------------------

\begin{table}[ht]
\centering
\begin{tabular}{l|r|r|r|r}
\parbox{0.2\textwidth}{\centering Linker} &
\parbox{0.2\textwidth}{\centering Comparisons} &
\parbox{0.17\textwidth}{\centering Precision} &
\parbox{0.17\textwidth}{\centering Recall} &
\parbox{0.17\textwidth}{\centering F1 Measure} \\ \hline \hline
Brute Force        & 1,677,025                 & 0.84      & 0.57   & 0.68       \\ \hline
M-tree              &  902,693                  & 0.84      & 0.57   & 0.68       \\ \hline
LSH-2-2            &  192,199	& 0.95	& 0.47	& 0.63       \\
LSH-5-2            &  342,849	& 0.91	& 0.55	& 0.69       \\
LSH-10-2           &  513,947	& 0.88	& 0.57	& 0.69       \\
LSH-2-5            &  14,329	& 0.99	& 0.28	& 0.43      \\
LSH-5-5            &  22,057	& 0.99	& 0.36	& 0.53       \\
LSH-10-5           &  26,167	& 0.98	& 0.4	& 0.57      \\
LSH-2-10           &  4,711	& 1	& 0.15	& 0.27    \\
LSH-5-10           &  6,501	& 1	& 0.19	& 0.32      \\
LSH-10-10          & 10,627  	& 0.99	& 0.27& 	0.43       \\  \hline
Block-year         &  115,893                  & 0.99      & 0.35   & 0.51       \\
Block-authors      &   11,039                   & 0.94      & 0.16   & 0.28       \\
Block-title        &   27,407                   & 0.95      & 0.42   & 0.58       \\
Block-venue        &   36,647                   & 0.85      & 0.29   & 0.44       \\
Block-location     & 1,009,957                 & 0.83      & 0.43   & 0.57       \\
Block-publisher    &  833,079                  & 0.85      & 0.44   & 0.58       \\
Block-combined     & 1,214,269                 & 0.84      & 0.56   & 0.67       \\ \hline
\end{tabular}
\caption{Linkage quality on Cora dataset with distance threshold = 70
\label{comparison-of-results-cora}
}
\end{table}

%----------------------------------------------------------------------

\subsection{Demographic Dataset Results}

@@

%----------------------------------------------------------------------

\begin{table}[ht]
\centering
\begin{tabular}{l|r|r|r|r}
\parbox{0.2\textwidth}{\centering Linker} &
\parbox{0.2\textwidth}{\centering Comparisons} &
\parbox{0.17\textwidth}{\centering Precision} &
\parbox{0.17\textwidth}{\centering Recall} &
\parbox{0.17\textwidth}{\centering F1 Measure} \\ \hline \hline
M-tree    & 102,318,525                 & 0.65      & 0.46   & 0.54      \\ \hline

@@

%----------------------------------------------------------------------

\begin{table}[ht]
\centering
\begin{tabular}{l|r|r|r|r|r|r}
\parbox{0.2\textwidth}{\centering Linker} &
\parbox{0.2\textwidth}{\centering Comparisons} &
\parbox{0.17\textwidth}{\centering Precision} &
\parbox{0.17\textwidth}{\centering Recall} &
\parbox{0.17\textwidth}{\centering F1 Measure} \\ \hline \hline
M-tree     & 514,871,153                   & 0.76      & 0.45   & 0.57      \\ \hline
LSH-2-2   & 99,145,887                    & 0.81      & 0.16   & 0.27       \\
LSH-5-2   & 130,721,338                   & 0.79      & 0.23   & 0.36       \\
LSH-10-2  & 177,168,848                   & 0.79      & 0.36   & 0.49       \\
LSH-5-5   & 239,368                      & 0.84      & 0.01   & 0.02       \\
LSH-10-5  & 855,431                      & 0.87      & 0.02   & 0.03 \\ \hline
\end{tabular}
\caption{Linkage quality on Kilmarnock dataset with distance threshold = 2
\label{comparison-of-results-demography-kili}
}
\end{table}

%----------------------------------------------------------------------

@@

However, this paper also highlights some unexpected results. Firstly, the good quality results yielded by both traditional blocking and LSH are partly due to the fact that (in the limit) they tend towards brute-force as the number of records in the block increase. A second unexpected result is that shown in \Cref{comparison-of-results-cora}, namely that LSH can yield higher precision than is achieved by a complete method such as M-tree, as was discussed in Sect.~\ref{sec-approach}.
